Using Scipion for Stream Image Processing at Electron Microscopy Facilities

### 2018-04-18 0.abstract.tex
from home Marta (swedden) comments

Three dimensional electron microscopy (3DEM) is becoming a very data-intensive field in which vast amounts of experimental images are acquired at high speed. To manage such large-scale projects, we had previously developed  a modular workflow system called \scipion \citep{delarosatrevin2016}. We present here a major extension of \scipion that allows stream processing of EM images while the data is being acquired. This approach helps to detect problems at early stages, saves computing time and provides users with an in deep evaluation of the data quality before the acquisition is finished. Different EM facilities have 
implemented their stream processing solution using custom-made scripts, but this script based approach is not flexible and present a slow response time to new feature requests. The advantages of our proposed solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or with a different data set; (3) monitorization: there are utilities that constantly check how the execution of the algorithms is going on and inform about possible anomalies; (4) reports: we have developed several graphical interfaces that are updated periodically producing a summary (e.g, CTF defocus values, system load, etc). This summary may be generated in HTML format and  published though a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided. At present, \scipion has been installed and is in production mode in seven Cryo-EM facilities and it is being considered in many more.

\textbf{Keywords:} electron microscopy

>>>>

Three dimensional electron microscopy is becoming a very data-intensive field in which vast amounts of experimental images are acquired at high speed. To manage such large-scale projects, we had previously developed  a modular workflow system called \scipion \citep{delarosatrevin2016}. We present here a major extension of \scipion that allows stream processing of EM images while the data is being acquired. This approach helps to detect problems at early stages, saves computing time and provides users with a detailed evaluation of the data quality before the acquisition is finished. Different EM facilities have 
implemented their stream processing solution using custom-made scripts, but this script based approach is not flexible and present a slow response time to new feature requests. The advantages of our proposed solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or with a different data set; (3) monitorization: there are utilities that constantly check how the execution of the algorithms is going on and inform about possible anomalies; (4) reports: we have developed several graphical interfaces that are updated periodically producing a summary (e.g, CTF defocus values, system load, etc). This summary may be generated in HTML format and  published though a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided. At present, \scipion has been installed and is in production mode in seven Cryo-EM facilities and it is being considered in many more.

\textbf{Keywords:} electron microscopy, streaming, image processing, live processing, high throughput

### 2018-04-18 1.introduction.tex
to home
from home Marta (swedden) comments

3D electron microscopy (3DEM)  can provide rich information about structural characteristics of macromolecular complexes. The field is under a major transformation due to the arrival of better microscopes, new large area direct electron detectors and automatization \citep{kuhlbrandt2014a, Kuhlbrandt2014b}. These improvements make possible that a single microscope generates high quality data sets on the order of terabytes per day \citep{Saibil2015} while working for several days without interruption. Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of a experienced microscopist and; (3) the need for an effective data and metadata transfer in an increasingly distributed and collaborative environment. %This situation strongly demands the application of techniques that are highly automated to allow high throughput.

@@

For most cases, the future of projects that require high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff.  \scipion has been redesigned with this scenario in mind. % in which the early detection of issues may save a data acquisition shift which otherwise needs to be repeated after a new application that may be granted many months later. 
In this way, \scipion starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and execute a first user-tailored image processing workflow that provides users with an accurate measurement of the acquired data quality.

>>>>

3D electron microscopy (3DEM)  can provide rich information about structural characteristics of macromolecular complexes. The field is under a major transformation due to the arrival of better microscopes, new large area direct electron detectors and automation \citep{kuhlbrandt2014a, Kuhlbrandt2014b}. These improvements make possible that a single microscope generates high quality data sets on the order of terabytes per day \citep{Saibil2015} while working for several days without interruption. Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of a experienced microscopist and; (3) the need for an effective data and metadata transfer in an increasingly distributed and collaborative environment. %This situation strongly demands the application of techniques that are highly automated to allow high throughput.

@@

For most cases, the future of projects that require high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff.  With this scenario in mind, we have redesigned \scipion which is a software framework for integrating several 3DEM software packages through a workflow-based approach \citep{delarosatrevin2016}. % in which the early detection of issues may save a data acquisition shift which otherwise needs to be repeated after a new application that may be granted many months later. 
In this way, \scipion starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and execute a first user-tailored image processing workflow that provides users with an accurate estimation of the acquired data quality.

### 2018-04-18 3.workflow_cnb.tex
from home Marta (swedden) comments

At the beginning CNB networking setup was a quite straight forward one, movies where stored in the  same machine that ran \scipion and processed data was saved to external USB hard disks by this same machine. The situation changed when the new Falcon-III detector was installed. After that, data was produced at sustained speeds as high as 90 MB/sec (when acquiring four movies by hole in lineal mode) and the main problem became to move smoothly the data through the entire pipeline from the microscope to the final user laboratory. In the following we describe such a pipeline.

@@

Although CNB staff try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processing data, this is the preferred option for most users.
USB 3.0 drives can theoretically handle data flows of 300 MB/s, but CNB experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s. In most cases, 50 MB/s is enough to transfer the data before the following microscope shift starts. Nevertheless, real-world is tough and it happens that two, or even three,  users try to download their data simultaneously. Fortunately, output-NAS works as front-end and isolates the data acquiring and processing infrastructure from the user downloading. Network users access to output-NAS using a chrooted common account. This account only allows rsync connections and make impossible to retrieve data if the name of the project to be downloaded is unknown. In order to speed up transfer weak ciphers (as \emph{arcfour}) has been activated in output-NAS.

@@

\emadmin was born as a script that created a normalized tree of directories  in which to store the movies acquired by the microscope, helped users to choose between a few predetermined image processing workflows and finally launched \scipion. At present it has grown in a client-server application that in addition to the old script capacities it has incorporate the following functions:

>>>>

At the beginning CNB networking setup was a quite straight forward one, movies where stored in the  same machine that ran \scipion and processed data were saved to external USB hard disks by this same machine. The situation changed when the new Falcon-III detector was installed. After that, data were produced at sustained speeds as high as 90 MB/sec (when acquiring four movies by hole in lineal mode) and the main problem became to move smoothly the data through the entire pipeline from the microscope to the final user laboratory. In the following we describe such a pipeline.

@@

Although CNB staff try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processed data, this is the preferred option for most users.
USB 3.0 drives can theoretically handle data flows of 300 MB/s, but CNB experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s. In most cases, 50 MB/s is enough to transfer the data before the following microscope shift starts. Nevertheless, real-world is tough and it happens that two, or even three,  users try to download their data simultaneously. Fortunately, output-NAS works as front-end and isolates the data acquiring and processing infrastructure from the user downloading. Network users access to output-NAS using a chrooted common account. This account only allows rsync connections and make impossible to retrieve data if the name of the project to be downloaded is unknown. In order to speed up transfer weak ciphers (as \emph{arcfour}) has been activated in output-NAS.

@@

\emadmin was born as a script that created a normalized tree of directories  
where movies acquired by the microscope could be stored, helped users to choose between a few predetermined image processing workflows and finally launched \scipion. At present it has grown in a client-server application that in addition to the old script capacities has incorporated the following functions:

### 2018-04-18 3.workflow_diamond.tex
from home Marta (swedden) comments

Diamond IT infrastructure is complex and requires to integrate a large collection of heterogeneous systems. In this environment a major problem is to achieve interaction and data exchange between diverse applications. One of the solutions to this problem has been to implement \emph{indirect communication} through a message queuing system called \emph{ActiveQ}. Simply said; a message queuing system is a software where queues can be defined, applications may connect to the queue and send or retrieve  messages. A message can include any kind of information, for example, about a process/task that should start on another application (that could be on another server). %The message queuing software stores the messages until a receiving application connects and takes a message off the queue. The receiving application then processes the message in an appropriate manner.

>>>>

Diamond IT infrastructure is complex and requires to integrate a large collection of heterogeneous systems. In this environment a major problem is to achieve interaction and data exchange between diverse applications. One of the solutions to this problem has been to implement \emph{indirect communication} through a message queuing system called \emph{ActiveQ}. A message queue is a mechanism
that allows a sender process and a receiver process to exchange information; instead of connecting directly, the sender posts messages in a queue and the receiver, when ready, takes these messages from the queue and process them appropriately.

%In this environment any piece of information or data that needs to be interchanged among  two applications will be placed in a queue by the sender application until the destination application explicitly asks for it.

%Simply said; a message queuing system is a software where queues can be defined, applications may connect to the queue and post or retrieve  messages. A message can include any kind of information, for example, about a process/task that should start on another application (that could be on another server). %The message queuing software stores the messages until a receiving application connects and takes a message off the queue. The receiving application then processes the message in an appropriate manner.

### 2018-04-18 5.conclusions.tex
from home Marta (swedden) comments

 In a workflow approach, users (including microscope staff) can reasonably be expected to edit the paths. They can potentially play with the pipeline in ways that aren't possible when having a programmer write code for them. Furthermore, not having the processing path hardcoded in the program means you don't have to recompile, so changes can be quickly tested   even at run time. The \scipion workflow approach provides provenance, that is,  for any created object (micrograph, CTFs, class, etc) \scipion stores metadata such as which programs, which parameters, and which exact path was followed to create any output data.  In this way,  real tracking and reproducibility are provided.

>>>>

 In a workflow approach, users (including microscope staff) can reasonably be expected to edit the paths. They can potentially play with the pipeline in ways that are not possible when having a programmer writing the code for them. Furthermore, not having the processing path hardcoded in the program means you do not have to recompile, so changes can be quickly tested   even at run time. The \scipion workflow approach provides provenance, that is,  for any created object (micrograph, CTFs, class, etc) \scipion stores metadata such as which programs, which parameters, and which exact path was followed to create any output data.  In this way,  real tracking and reproducibility are provided.

### 2018-04-18 scipionbox.tex
carazo's review
to home
from home Marta (swedden) comments

\begin{document}
\title{Using Scipion for Stream Image Processing at EM Facilities}
Authors to be sorted later
\author{C.O.S. Sorzano$^{(a)}$,
A. Jim\'{e}nez$^{(a)}$,
R. Melero$^{(a)}$,\\
M. Mart\'{i}nez$^{(a)}$,
J. G\'{o}mez-Blanco$^{(a)}$,
L. del Cano$^{(a)}$,\\
T. Majtner$^{(a)}$,
E. Ram\'{i}rez$^{(a)}$,
P. Conesa$^{(a)}$,\\
J.L. Vilas$^{(a)}$,
R. Marabini$^{(a)}$,
Y. Rancel$^{(a)}$,
J.M. Carazo$^{(a)}$,\\
J.M. de la Rosa-Trev\'{i}n$^{(c)}$,
M. Carroni$^{(c)}$,
S. Fleichman $^{(c)}$,\\
E. Lindahl$^{(c,d)}$
A.W. Ashton$^{(e)}$,
M. Basham$^{(e)}$,\\
D.K. Clare$^{(e)}$,
K. Savage$^{(e)}$,
C.A. Siebert$^{(e)}$,\\
D. Maluenda$^{(a)}$
        }

@@

\lyxaddress{{\small \vspace{-0.5cm}$^{(a)}$Escuela Polit\'{e}cnica Superior, Universidad Aut\'{o}noma de Madrid, 28049 Cantoblanco, Madrid, Spain.}}
\lyxaddress{{\small$^{(b)}$ Biocomputing Unit, National Center for Biotechnology (CSIC), C/ Darwin, 3, Campus Universidad Aut\'{o}noma, 28049 Cantoblanco, Madrid, Spain.}}
\lyxaddress{{\small$^{(c)}$ Department of Biochemistry and Biophysics, Science for Life Laboratory, Stockholm University, Stockholm, Sweden}} 
\lyxaddress{{\small$^{(d)}$ Swedish e-Science Research Center, KTH Royal Institute of Technology, Stockholm, Sweden}}
\lyxaddress{{\small$^{(e)}$ Diamond Light Source, Harwell Science and Innovation Campus, Didcot OX11 0DE, England}}
%This document provides a preliminary analysis of the data collected by the CTF Challenge (http://i2pc.cnb.csic.es/3dembenchmark/LoadHome.htm). Figures/tables/plots/etc are located at the document's end. 
\setcounter{figure}{2}%I do not know why but figure counter starts with -1
\textbf{Corresponding Author:} %Roberto Marabini, Ph= 34 91 5854510, email: roberto@cnb.csic.es\\
%Escuela Polit\'{e}cnica Superior\\
%Universidad Aut\'{o}noma de Madrid\\
%28049 Cantoblanco\\
%Madrid\\
%Spain

>>>>

\begin{document}
\title{Using Scipion for Stream Image Processing at Electron Microscopy Facilities}
Authors to be sorted later
\author{C.O.S. Sorzano$^{(a)}$,
A. Jim\'{e}nez$^{(a)}$,
R. Melero$^{(a)}$,\\
M. Mart\'{i}nez$^{(a)}$,
J. G\'{o}mez-Blanco$^{(a)}$,
L. del Cano$^{(a)}$,\\
T. Majtner$^{(a)}$,
E. Ram\'{i}rez$^{(a)}$,
P. Conesa$^{(a)}$,\\
J.L. Vilas$^{(a)}$,
R. Marabini$^{(a)}$,
Y. Rancel$^{(a)}$,
J.M. Carazo$^{(a)}$,\\
J.M. de la Rosa-Trev\'{i}n$^{(c)}$,
M. Carroni$^{(c)}$,
S. Fleichman $^{(c)}$,\\
E. Lindahl$^{(c,d)}$
A.W. Ashton$^{(e)}$,
M. Basham$^{(e)}$,\\
D.K. Clare$^{(e)}$,
K. Savage$^{(e)}$,
C.A. Siebert$^{(e)}$,\\
D. Maluenda$^{(a)}$
        }

@@

\lyxaddress{{\small \vspace{-0.5cm}$^{(a)}$Escuela Polit\'{e}cnica Superior, Universidad Aut\'{o}noma de Madrid, 28049 Cantoblanco, Madrid, Spain.}}
\lyxaddress{{\small$^{(b)}$ Biocomputing Unit, National Center for Biotechnology (CSIC), C/ Darwin, 3, Campus Universidad Aut\'{o}noma, 28049 Cantoblanco, Madrid, Spain.}}
\lyxaddress{{\small$^{(c)}$ Department of Biochemistry and Biophysics, Science for Life Laboratory, Stockholm University, Stockholm, Sweden}} 
\lyxaddress{{\small$^{(d)}$ Swedish e-Science Research Center, KTH Royal Institute of Technology, Stockholm, Sweden}}
\lyxaddress{{\small$^{(e)}$ Diamond Light Source, Harwell Science and Innovation Campus, Didcot OX11 0DE, England}}
%This document provides a preliminary analysis of the data collected by the CTF Challenge (http://i2pc.cnb.csic.es/3dembenchmark/LoadHome.htm). Figures/tables/plots/etc are located at the document's end. 
\setcounter{figure}{2}%I do not know why but figure counter starts with -1
\textbf{Corresponding Author: Roberto Marabini, Ph = +34 91 5854510, email: roberto@cnb.csic.es\\} %Roberto Marabini, Ph= 34 91 5854510, email: roberto@cnb.csic.es\\
%Escuela Polit\'{e}cnica Superior\\
%Universidad Aut\'{o}noma de Madrid\\
%28049 Cantoblanco\\
%Madrid\\
%Spain
